---
title: "STAT525 HW2"
author: "Yuxuan Zhang"
date: "2/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1
Design a rejection sampling algorithm for generating samples from a distribution with the following pdf:
$$\pi(x) = \frac{3}{2}x^3 +\frac{11}{8}x^2+\frac{1}{6}x+\frac{1}{12}, 0\leq x\leq 1 $$
$$\pi(x) \leq cg(x)$$
$$g(x) = Uniform[0,1], c = \max_{[0,1]}\pi(x)= \pi(1) = \frac{25}{8}$$
```{r}
pi = function(x) {3/2*x^3 + 11/8*x^2 + 1/6*x + 1/12}
c = 25/8
i = 1
sample = c(rep(0,1000))
while (i <= 1000) {
  x = runif(1)
  u = runif(1)
  if (u <= pi(x)/c) {
    sample[i] = x
    i = i + 1
  }
}
# Estimate of E(x^2)
mean(sample^2)
# standard error of the estimate
sqrt(var(sample^2)/1000)
```

### 2
$$\pi(x)\propto \{[sin(8x)]^2 + 2[cos(3x)]^4 + 1]\}e^{-x}, 0\leq x \leq \infty$$
$$\pi(x)\propto l(x),\quad  l(x) \leq cg(x)$$
$$g(x) = e^{-x}, 0\leq x \leq \infty ; \quad c = \max([sin(8x)]^2 + 2[cos(3x)]^4 + 1) = 4$$
$$u \leq \frac{l(x)}{cg(x)} = \frac{[sin(8x)]^2 + 2[cos(3x)]^4 + 1}{4}$$
```{r}
pi = function(x) {(sin(x)^2 + 2*cos(3*x)^4 + 1)}
c = 4
i = 1
sample = c(rep(0,1000))
while (i <= 1000) {
  x = rexp(1)
  u = runif(1)
  if (u <= pi(x)/c) {
    sample[i] = x
    i = i + 1
  }
}
# Estimate of the mean of \pi(x)
mean(sample)
# standard error of the estimate
sqrt(var(sample)/1000)
```
### 3
Let $X_1,...,X_n$ be i.i.d from $N(0,1)$ where $\theta$ is the unknown parameter. In Bayesian inference, we may put a Cauchy prior distribution on $\theta$ and Bayes estimate of $\theta$ is the mean of posterior distribution
$$\pi(\theta|x_1,...x_n) \propto \frac{1}{\pi(1+\theta^2)}\prod_{i=1}^{n}\frac{1}{\sqrt{2\pi}}e^{-(x_i-\theta)^2/2}$$
$$\pi(x)\propto l(x),\quad  l(x) \leq cg(x)$$
\begin{equation}
\begin{split}
l(\theta) &= \frac{1}{\pi(1+\theta^2)}(2\pi)^{-5/2}exp(-\frac{1}{2}[\sum(x_i-\theta)^2]) \\ 
     &= \frac{1}{\pi(1+\theta^2)}(2\pi)^{-5/2}exp(-\frac{1}{2}[\sum x_i^2+5\theta^2-2\theta\sum x_i]) \\
     &= \frac{1}{\pi(1+\theta^2)}(2\pi)^{-5/2}exp(-\frac{1}{2}[5(\theta^2-\frac{2\sum x_i}{5}\theta + \frac{\sum x_i^2}{5})]) \\
     &= \frac{1}{\pi(1+\theta^2)}(2\pi)^{-5/2}exp(-\frac{5}{2}[(\theta-\frac{\sum x_i}{5})^2 + \frac{\sum x_i^2}{5} - (\frac{\sum x_i}{5})^2]) \\
     &\leq \frac{1}{\pi}\frac{1}{\sqrt{5}} \frac{\sqrt{5}}{\sqrt{2\pi}}exp(-\frac{5}{2}(\theta-\frac{\sum x_i}{5})^2) \frac{1}{4\pi^2}exp(-\frac{\sum x_i^2}{2} + \frac{(\sum x_i)^2}{10}) \\
     &=\frac{1}{\sqrt{80}\pi^3}exp(-\frac{\sum x_i^2}{2} + \frac{(\sum x_i)^2}{10})N(\frac{\sum x_i}{5}, \frac{1}{5}) \\
     &= cg(\theta)
\end{split}
\end{equation}

Instrucmental distribution is $N(\frac{\sum x_i}{5}, \frac{1}{5})$

$$\frac{\pi(\theta)}{cg(\theta)}=\frac{\frac{1}{\pi(1+\theta^2)}}{\frac{1}{\sqrt{80}\pi^3}exp(-\frac{\sum x_i^2}{2} + \frac{(\sum x_i)^2}{10})} = \frac{\pi^2\sqrt{80}}{(1+\theta^2)exp(-\frac{\sum x_i^2}{2} + \frac{(\sum x_i)^2}{10})}$$

```{r}
x = c(1.6, 0.6, -0.7, 1.1, 0.8)

i = 1
N = 0
sample = c(rep(0,1000))
while (i <= 1000) {
  N = N + 1
  theta = rnorm(1, mean=sum(x)/5, sd=sqrt(1/5))
  if(runif(1) <= 1/(1+theta^2)){
    sample[i] = theta
    i=i+1
  }
}
# the number of samples I need to generate from the instrumental distribution in order to have 1000 accepted samples.
N
# estimated acceptance of rate
1000/N
# the mean of pi(theta|X)
mean(sample)
# standard error of my estimate
sqrt(var(sample)/1000)
```